---
title: "Homework 3"
author: Nikhita Damaraju
output: github_document
---

This is my solution to Homework 3.

Library imports  ###HIDE OUTPUT HERE###

```{r} 
library(p8105.datasets)
library(ggplot2)
library(tidyverse)
```

# Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are the level of items in orders by the user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. Products are in aisles and aisles are in departments.

## Number of aisles and the ones most items are from

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

## Plot of number of items ordered in each aisles with more than 10k items 

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Table showing most popular items 

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

## Table showing mean hour of the day

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>%
  knitr::kable()
```

# Section 2

I can take the mean of the sample, too!
